A Tale of Two Transects

### 2019-12-19 writing/transect_paper.Rmd
start putting results figures in manuscript

---
title: "A Tale of Two Transects"
subtitle: "Also Two Sequencing Technologies, and Two Amplicons"
author: 
  - Brendan Furneaux^1^,
    Roël Houdanon^2^,
    Mohammad Bahram^1^,
    Anna Rosling^3^,
    Nourou S. Yorou^2^,
    Martin Ryberg^1^
  - ^1^Program in Systematic Biology,
    Department of Organismal Biology,
    Uppsala University,
    Uppsala, Sweden
  - ^2^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
    LEB,
    University of Parakou,
    Parakou, Benin
  - ^3^Program in Evolutionary Biology,
    Department of Ecology and Genetics,
    Uppsala University,
    Uppsala, Sweden
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography:
  - "all.bib"
  - "R.bib"
csl: fungal-ecology.csl
panflute-filters: figurefilter
panflute-path: '../../python'
output:
  bookdown::pdf_document2:
    fig_crop: no
    keep_tex: yes
    latex_engine: lualatex
    includes:
      in_header: preamble.tex
  bookdown::word_document2: 
    keep_md: yes
---

```{r setup, include=FALSE}
library("knitr")
library("tikzDevice")
tryCatch(
  tinytex::tinytex_root(),
  error = function(e) tinytex::install_tinytex()
  )
knitr::opts_chunk$set(dev="tikz",
                      # dev.args = list(pointsize=32),
                      echo=FALSE,
                      results='asis',
                      fig.show='asis',
                      external=TRUE)
options(tikzDefaultEngine = "luatex",
        tikzLualatexPackages = c(options("tikzXelatexPackages")$tikzXelatexPackages,
                                "\\usepackage[version=4]{mhchem}\n",
                                "\\usepackage{siunitx}\n",
                                "\\DeclareSIUnit\\year{a}\n"))

base.dir <- regmatches(getwd(), regexpr(".*oueme-fungi-transect", getwd()))
if (opts_chunk$get("dev") == "tikz") {
  # styles$Expression <- styles$LaTeX.Expression
  # styles$Name <- styles$LaTeX.Name
  # styles$Short.Name <- styles$LaTeX.Short.Name
  # styles$Unit <- paste0("\\si{", styles$LaTeX.Unit, "}")
}
write_bib(c("base", "vegan", "permute"), file = "R.bib", tweak = TRUE)
```
# Introduction

@@

- Clustering vs. phylogenetic methods
    - Clustering-based OTU approach may both "clump" and "split" species.
    - Distance-based clustering conflates intra-species variation and sequencing error
    - Denoising (DADA2 [@callahan2017], Deblur [@amir2017], UNOISE2 [@edgar2016]) attempts to control for sequencing error while leaving intra-species variation.
    - Phylogenetic community distance measures like UNIFRAC [@wong2016; look for earlier reference] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.

@@

Amplicons were purified using SPRI beads [Vesterinen et al 2016] and quantified fluorometrically using PicoGreen dsDNA (Quant-iT) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan) according to the manufacturer's protocol.
100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences RS II sequencer at the Uppsala Genome Center (UGS) at SciLifeLab in Uppsala, Sweden.
Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.

@@

  - Two workflows for short reads:
    - Mohammad's workflow in Pipecraft
    - Brendan's workflow using Snakemake/Drake

@@

Raw reads were divided into regions/domains by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
First, the 5.8S rRNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013].
This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013; data not shown].
Each read was then aligned to a CM based on a modification of the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017, @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
The reference line in the CM alignment for each read were then used to split the reads into alternating more-conserved and less-conserved regionss:
ITS1, 5.8S, ITS2, LSU1, D1, LSU2, D2, LSU3, D3, LSU4,
where the four "LSU" regions represent the conserved regions of the 28S rRNA flanking the "D" regions.
For short amplicons, only 5.8S, ITS2, and LSU1 were extracted.
No attempt was made to remove the approximately 12 bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx.
The process of extracting the regions is automated in the new R package `LSUx`, available on github, CRAN, and bioconda (**but not yet**).

Each of the extracted regions was independently filtered for length (table **X**) and a maximum of three expected errors.
Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")`.
The error model for DADA2 denoising was fit using the 5.8S RNA region for long amplicons, and using the entire read for short amplicons.
Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, different sequencing technologies).
Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.

@@

The tree was rooted outside the fungi by using the most abundant ASV which was confidently assigned to the Chlorophyta by all 6 applicable taxonomic assignment methods.
Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
Inspection of the tree along with taxonomic assignments revealed that the genus *Tulasnella* (Basidiomycota), which is known to have accelerated evolution in the rDNA [@moncalvo2006], was nested within the Metazoa, so it was excluded from further analysis.
The kingdom Fungi was then identified as the minimal clade containing all remaining ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
ASVs falling outside this clade were excluded from further analysis.

@@

Ecological community dissimilarity matrices were calculated using methods "bray" (both long and short amplicons) and "unifrac" (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12 m,

@@

>>>>

---
title: "A Tale of Two Transects"
subtitle: "Also Two Sequencing Technologies, and Two Amplicons"
author: 
  - Brendan Furneaux^1^,
    Roël Houdanon^2^,
    Mohammad Bahram^1^,
    Anna Rosling^3^,
    Nourou S. Yorou^2^,
    Martin Ryberg^1^
  - ^1^Program in Systematic Biology,
    Department of Organismal Biology,
    Uppsala University,
    Uppsala, Sweden
  - ^2^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
    LEB,
    University of Parakou,
    Parakou, Benin
  - ^3^Program in Evolutionary Biology,
    Department of Ecology and Genetics,
    Uppsala University,
    Uppsala, Sweden
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography:
  - "all.bib"
  - "R.bib"
csl: molecular-ecology.csl
output:
  bookdown::pdf_document2:
    fig_crop: yes
    keep_tex: yes
    latex_engine: lualatex
    includes:
      in_header: preamble.tex
  bookdown::word_document2: 
    keep_md: yes
    fig_crop: yes
---

```{r setup, include=FALSE}
library("knitr")
library("tikzDevice")
library("ggplot2")
library("tidyverse")
library("drake")
library("here")
tryCatch(
  tinytex::tinytex_root(),
  error = function(e) tinytex::install_tinytex()
  )
knitr::opts_chunk$set(
  dev = "pdf",
  # dev.args = list(pointsize=32),
  echo = FALSE,
  # results = 'asis',
  # fig.show = 'asis',
  fig.width = 6,
  fig.height = 4.5,
  message = FALSE,
  warning = FALSE,
  external = TRUE
)
options(
  tikzDefaultEngine = "luatex",
  tikzXelatexPackages = c(options("tikzXelatexPackages")$tikzXelatexPackages,
                          "\\usepackage[version=4]{mhchem}\n",
                          "\\usepackage{siunitx}\n",
                          "\\DeclareSIUnit\\year{a}\n",
                          "\\usepackage{fontspec}"),
  tikzMetricsDictionary = here("writing/transect_paper-tikzDictionary")
)

base.dir <- regmatches(getwd(), regexpr(".*oueme-fungi-transect", getwd()))
if (opts_chunk$get("dev") == "tikz") {
  # styles$Expression <- styles$LaTeX.Expression
  # styles$Name <- styles$LaTeX.Name
  # styles$Short.Name <- styles$LaTeX.Short.Name
  # styles$Unit <- paste0("\\si{", styles$LaTeX.Unit, "}")
}
write_bib(c("base", "vegan", "permute"), file = "R.bib", tweak = TRUE)
cache = drake_cache(here(".light"))
light_plan <- readRDS(here("data/plan/drake_light.rds"))
theme_set(theme_bw())
```
# Introduction

@@

- Clustering vs. phylogenetic methods
    - Clustering-based OTU approach may both "clump" and "split" species.
    - Distance-based clustering conflates intra-species variation and sequencing error
    - Denoising (DADA2 [@callahan2017], Deblur [@amir2017], UNOISE2 [@edgar2016]) attempts to control for sequencing error while leaving intra-species variation.
    - Phylogenetic community distance measures like UNIFRAC [@wong2016, look for earlier reference] are relatively insensitive to species/OTU delimitation, but require a phylogenetic tree.

@@

Each PCR run also included a blank sample and a positive control consisting of freshly extracted DNA from a commercially purchased fruitbody of *Agaricus bisporus*.

Amplicons were purified using SPRI beads [@vesterinen2016] and quantified fluorometrically using PicoGreen dsDNA (Quant-iT) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan) according to the manufacturer's protocol.
100 ng of DNA from each sample (or the total PCR product if less than 100 ng) was pooled into two libraries each for long and short amplicons.
Each library was sequenced using Single Molecule Real Time (SMRT) sequencing on a Pacific Biosciences RS II sequencer at the Uppsala Genome Center (UGS) at SciLifeLab in Uppsala, Sweden.
Short amplicon libraries were sequenced on two SMRT cells each, while long amplicon libraries were sequenced on four SMRT cells each.

@@

@@

Raw reads were divided into regions/domains by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
First, the 5.8S rRNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013].
This approach was able to identify both the 5' and 3' ends of the 5.8S rRNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].
Each read was then aligned to a CM based on a modification of the fungal 28S RNA seed alignment from the Ribosomal Data Project (RDP) release 11.5 [@glockner2017; @cole2014] using `cmalign` from Infernal 1.1.2 [@nawrocki2013].
The reference line in the CM alignment for each read were then used to split the reads into alternating more-conserved and less-conserved regionss:
ITS1, 5.8S, ITS2, LSU1, D1, LSU2, D2, LSU3, D3, LSU4,
where the four "LSU" regions represent the conserved regions of the 28S rRNA flanking the "D" regions.
For short amplicons, only 5.8S, ITS2, and LSU1 were extracted.
No attempt was made to remove the approximately 12 bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx.
The process of extracting the regions is automated in the new R package `LSUx`, available on github, CRAN, and bioconda (**but not yet**).

```{r region-length}
datasets <- read_csv(here("config/datasets.csv"), col_types = "cccicccccicc")
regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
regions %>%
  select(Region = region, "Minimum length" = min_length, "Maximum length" = max_length) %>%
  right_join(tibble(Region = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2", "LSU3", "D3", "LSU4"))) %>%
  mutate_at("Region", str_replace_all, "_", ".") %>%
  knitr::kable(
    booktabs = TRUE,
    caption = "Minimum and maximum allowed lengths for each extracted region."
  )
```

Each of the extracted regions was independently filtered for length (Table \@ref(tab:region-length)) and a maximum of three expected errors.
Sequences were then dereplicated and denoised into amplicon sequencing variants (ASVs) using DADA2 version `r packageVersion("dada2")`.
The error model for DADA2 denoising was fit using the 5.8S RNA region for long amplicons, and using the entire read for short amplicons.
Independent error models were fit for each sequencing run (i.e., long *vs.* short amplicons, different sequencing technologies).
Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.

@@

The tree was rooted outside the Fungi by using the most abundant ASV which was confidently assigned to a non-fungal kingdom by all 6 applicable taxonomic assignment methods.
Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset.
Inspection of the tree along with taxonomic assignments revealed that the genus *Tulasnella* (Basidiomycota), which is known to have accelerated evolution in the rDNA [@moncalvo2006], was nested within the Metazoa, so it was excluded from further analysis.
The kingdom Fungi was then identified as the minimal clade containing all remaining ASVs which were confidently identified (consensus of at least 6 of 9 assignments) to a fungal phylum.
ASVs falling outside this clade were excluded from further analysis.

@@

Ecological community dissimilarity matrices were calculated using the species based Bray-Curtis metric (both long and short amplicons) and the phylogenetically based weighted UNIFRAC method (only long amplicons) in `phyloseq` version `r packageVersion("phyloseq")`.
Each of these distance matrices was used to calculate a Mantel correlogram for distances of 0--12 m.
Separate correlograms were drawn for samples taken during the same year, and samples separated in time by one year, in order to assess the degree to which the soil community changes over the course of one year.

Additionally, dissimilarities were plotted as a function of spatial distance up to 25 m and fit using an exponential variogram function using `gstat` version `r packageVersion("gstat")` using weighting .
The mean dissimilarity between samples taken in different transects (distance $\approx$ 30 km) was also included as a single point for each year for the purposes of fitting the variogram function.

@@

```{r read-counts}

- *Agaricus bisporus* from positive control was most abundant ASV in all sequencing runs
  - only one ASV in both PacBio runs
  - several ASVs in IonTorrent -- due to sequencing errors not controlled by DADA2 or intragenome variation only detected in deep sequencing?

(ref:readcap) ASVs/OTUs which were detected by one sequencing strategy but not the other are plotted as tick marks along the axes. Dashed line represents a constant ratio of read numbers. The blue line is a LOESS smooth of the data, with associated uncertainty in grey shading.

```{r pbl-vs-its, fig.cap="Comparison between read number for PacBio RSII long reads and Ion Proton S5 reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
loadd(big_seq_table_ITS2, cache = cache)
asv_table <- big_seq_table_ITS2 %>%
  tibble::as_tibble(rownames = "filename") %>%
  tidyr::gather(key = "seq", value = "reads", -1) %>%
  dplyr::filter(reads >= 1) %>%
  tidyr::extract(col = "filename",
                 into = c("seq.run", "plate", "well", "dir", "region"),
                 regex = "([a-z]+_\\d+)_(\\d+)_([A-H]1?[0-9])([fr]?)_([:alnum:]+).+") %>%
  dplyr::group_by(seq.run, seq) %>%
  dplyr::summarize(reads = sum(reads)) %>%
  tidyr::spread(key = seq.run, value = reads, fill = 0)

otu <- read_tsv(
  here("data/clusters/ITS2.table"),
  col_types = cols(
    .default = col_integer(),
    `#OTU ID` = col_character()
  )
) %>%
  column_to_rownames("#OTU ID") %>%
  as.matrix() %>%
  t() %>%
  as_tibble(rownames = "sample") %>%
  tidyr::extract(
    col = "sample",
    into = c("seq_run", "plate", "well"),
    regex = "([a-z]{2}_\\d{3})_(\\d{3})([A-H]1?\\d)"
  ) %>%
  tidyr::gather(key = "seq", value = "reads", -(1:3)) %>%
  dplyr::filter(reads >= 1) %>%
  dplyr::group_by(seq_run, seq) %>%
  dplyr::summarize(reads = sum(reads)) %>%
  tidyr::spread(key = seq_run, value = reads, fill = 0)

big_table <- bind_rows(
  mutate(asv_table, type = "ASV"),
  mutate(otu, type = "OTU")
) %>%
  select(-seq)

comparisons <- crossing(
  x = factor(datasets$seq_run, ordered = TRUE),
  y = factor(datasets$seq_run, ordered = TRUE),
  type = c("ASV", "OTU")
) %>%
  filter(y > x) %>%
  mutate_all(as.character) %>%
  mutate(
    cross = paste(x, y, sep = "-")
  ) %>%
  group_by(cross, type) %>%
  group_modify(function(d, g) {
    big_table %>%
      filter(type == g$type) %>%
      select(!!d$x, !!d$y) %>%
      filter(.data[[d$x]] > 0 & .data[[d$y]] > 0) %>%
      mutate_all(log) %>%
      lm() %>%
      broom::glance() %>%
      bind_cols(d, .)
    })

compR2 <- function(x, y, type) {
  format(filter(comparisons, x == !!x, y == !!y, type == !!type)$r.squared, digits = 2, drop0trailing = FALSE)
}

big_table %>%
  filter(is_057 > 0, pb_500 > 0) %>%
  ggplot(aes(x = is_057, y = pb_500)) +
  geom_point(alpha = 0.2, shape = 1) +
  scale_x_log10() +
  scale_y_log10(limits = c(1, 1e4)) +
  coord_equal() +
  xlab("Ion Torrent Short Reads") +
  ylab("PacBio Long Reads") +
  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
           data = big_table %>% filter(is_057 == 0, pb_500 > 0)) +
  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
           data = big_table %>% filter(pb_500 == 0, is_057 > 0)) +
  stat_smooth(method = "loess", formula = y ~ x) +
  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
                    mean(log10(is_057[is_057 >= 1])), slope = 1),
              linetype = 2L, alpha = 0.5) +
  facet_wrap(~type)
```

- Figure \@ref(fig:pbl-vs-its)
  - ASVs/OTUs are correlated between Ion Torrent and PacBio; 
  ASVs $R^2 = `r compR2("is_057", "pb_500", "ASV")`$; 
  OTUs $R^2 = `r compR2("is_057", "pb_500", "OTU")`$
  - "hook" due to sampling effects with rare species
  - lots of ASVs/OTUs detected in only one tech

- Figure \@ref(fig:pbs-vs-its)
  - very few PacBio OTUs undetected by Ion torrent from same library; only rare OTUS undetected by PacBio
  - $R^2 = `r compR2("is_057", "pb_483", "OTU")`$
  - sequencing technologies are comparable, but ASVs not so much
  - not such good match with ASVs -- suggests DADA2 doesn't perform equally

- Figure \@ref(fig:pbl-vs-pbs)
  - No "hook", because depth is similar
  - still lots of ASVs/OTUs detected by only one amplicon -- primer bias.

```{r pbs-vs-its, message=FALSE, fig.cap="Comparison between read number for PacBio RSII short reads and Ion Proton S5 reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
big_table %>% 
  filter(is_057 > 0, pb_483 > 0) %>%
  ggplot(aes(x = is_057, y = pb_483)) +
  geom_point(alpha = 0.2, shape = 1) +
  scale_x_log10() +
  scale_y_log10(limits = c(1, 1e4)) +
  xlab("Ion Torrent Short Reads") +
  ylab("PacBio Short Reads") +
  coord_equal() +
  geom_rug(aes(y = pb_483, x = 1), sides = "l", alpha = 0.2,
           data = big_table %>% filter(is_057 == 0, pb_483 > 0)) +
  geom_rug(aes(y = is_057), sides = "b", alpha = 0.2,
           data = big_table %>% filter(pb_483 == 0, is_057 > 0)) +
  stat_smooth(method = "loess", formula = y ~ x) +
  geom_abline(aes(intercept = mean(log10(pb_483[pb_483 >= 1])) -
                    mean(log10(is_057[is_057 >= 1])), slope = 1),
              linetype = 2L, alpha = 0.5) +
  facet_wrap(~type)

```{r pbl-vs-pbs, message=FALSE, fig.cap="Comparison between read number for PacBio RSII long and short amplicon reads, by ASV (left) and 95% OTU (right). (ref:readcap)"}
big_table %>% 
  filter(pb_483 > 0, pb_500 > 0) %>%
  ggplot(aes(x = pb_483, y = pb_500)) +
  geom_point(alpha = 0.2, shape = 1) +
  scale_x_log10() +
  scale_y_log10() +
  coord_equal() +
  xlab("PacBio Short Reads") +
  ylab("PacBio Long Reads") +
  geom_rug(aes(y = pb_500, x = 1), sides = "l", alpha = 0.2,
           data = big_table %>% filter(pb_483 == 0, pb_500 > 0)) +
  geom_rug(aes(y = pb_483), sides = "b", alpha = 0.2,
           data = big_table %>% filter(pb_500 == 0, pb_483 > 0)) +
  stat_smooth(method = "loess", formula = y ~ x) +
  geom_abline(aes(intercept = mean(log10(pb_500[pb_500 >= 1])) -
                    mean(log10(pb_483[pb_483 >= 1])), slope = 1),
              linetype = 2L, alpha = 0.5) +
  facet_wrap(~type)
```

(ref:correlog-cap) Mantel correlograms showing the correlation between Bray-Curtis (left two columnn) or weighted Unifrac (right column) dissimilarities and spatial distance, for short (left column) and long (right column) amplicons, and for the entire soil fungal community (top row) or only ECM taxa (bottom row).  Unifrac distance was not calculated for short amplicons because it requires that the sequences are placed on a phylogenetic tree.

```{r correlog-ecm, fig.cap="(ref:correlog-cap)"}

light_plan %>%
  filter(startsWith(target, "correlog")) %>%
  select(target) %>%
  separate(
    target,
    into = c("step", "guild", "metric", "tech", "amplicon", "tdist"),
    sep = "_",
    remove = FALSE
  ) %>%
  mutate(
    tdist = paste(tdist, "year"),
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
  ) %>%
  mutate_at("target", purrr::map, readd, character_only = TRUE, cache = cache) %>%
  mutate_at("target", purrr::map, "mantel.res") %>%
  mutate_at("target", purrr::map, as.data.frame) %>%
  unnest(target) %>%
  filter(complete.cases(.)) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3, size = 1.5) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(limits = c(-0.05, NA), name = "Mantel correlation") +
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
  scale_color_discrete(name = "Time lag") +
  xlab("Distance (m)") +
  facet_grid(guild ~ metric + amplicon) +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
```

(ref:variog) Spatial variogram fits for community dissimilarities. Each point represents one pair of samples taken at the same time.  The blue line and shaded area represent a LOESS smooth of the data and its associated confidence interval, respectively.  The red line is an exponential fit for the relationship between community dissimilarity and spatial distance.  Horizontal dashed lines represent the values of the fit function at zero distance ("nugget") and infinite distance ("sill").  The vertical dashed line is at the characteristic distance of the fit function ("range").

```{r variog, warning = FALSE, message = FALSE}
variog_data <- light_plan %>%
  filter(startsWith(target, "variofit")) %>%
  select(target, dist, dist_spatial) %>%
  separate("target", c("step", "guild", "metric", "tech", "amplicon"), sep = "_", remove = FALSE) %>%
  mutate(
    guild = plyr::mapvalues(guild, c("fungi", "ecm"), c("all Fungi", "ECM")),
    metric = plyr::mapvalues(metric, c("bray", "wunifrac"), c("Bray-Curtis", "Weighted UNIFRAC")),
    amplicon = factor(amplicon, levels = c("short", "long"), labels = c("short amplicon", "long amplicon"))
  ) %>%
  mutate_at(c("target", "dist", "dist_spatial"), lapply, readd, character_only = TRUE, cache = cache) %>%
  mutate_at(c("dist", "dist_spatial"), lapply, unclass) %>%
  mutate(sills = purrr::map(target, ~ cumsum(.$psill)),
         range = purrr::map_dbl(target, ~.$range[2]),
         vgline = purrr::map(target, gstat::variogramLine, maxdist = 25, n = 100))

variog_points <- unnest(variog_data, c("dist", "dist_spatial")) %>%
  filter(dist_spatial <= 20)
variog_empir <- variog_points %>%
  group_by(guild, metric, tech, amplicon, dist_spatial) %>%
  summarize_at("dist", mean)
variog_sills <- unnest(variog_data, "sills")
variog_fitline <- unnest(select(variog_data,-"dist"), "vgline")

variog_points %>%
  ggplot(aes(dist_spatial, dist)) +
  geom_hline(data = variog_sills, aes(yintercept = sills), linetype = "dashed", color = "gray50") +
  geom_vline(data = variog_data, aes(xintercept = range), linetype = "dashed", color = "gray50") +
  geom_point(shape = 1, alpha = 0.2) +
  geom_line(data = variog_empir, color = "blue", size = 1.5) +
  geom_line(data = variog_fitline,
            aes(dist, gamma), inherit.aes = FALSE, color = "red", size = 0.5) +
  scale_x_continuous(expand = expand_scale(0, 0), limits = c(0, 25), name = "Distance (m)") +
  ylab("Community dissimilarity") +
  facet_grid(guild ~ metric + amplicon)
